Report

### 2023-01-23 manuscript.Rmd

This simulation study validated the product Bayes factor as an alternative or complementary method to synthesize evidence across studies.
The PBF was benchmarked against other common methods for evidence synthesis.
This was done by calculating the accuracy, sensitivity and specificity of the methods in different plausible varying conditions.
During the simulation the hypotheses remained constant $H_i: \rho > .1$, $H_0: \rho = .1$.
To be able to calculate both sensitivity and specificity, half of the simulated conditions contained an effect, $\rho = .2$, while the other half did not, $\rho = .1$.
<!-- EB: Why did we not use H0:r=0 and Hi:r>0 as constant hypotheses? It seems more intuitive -->
Sensitivity and specificity can be interpreted as Type I error rate and statistical power respectively and are used interchangeably.
 <!-- Because Type 1 error rate and power seem to be key terms in the discussion, we may be able to include it in the introduction: 'How does the PBF compare to other state-of-the-art algorithms with regards to Type 1 error rate, power and overall accuracy?' -->
The effect of sample size per study, number of studies and reliability of measurement on the performance of the algorithms was investigated.

@@

With an increasing number of groups, slightly decreasing specificity was traded off for increasing sensitivity.<!--EB: I would argue that the decrease is negligible, 0.767 at k =2, 0.763 at k = 10 -->
With an increasing number of groups, specificity remained around .76 for PBF, while sensitivity increased to .87.
Again, the specificity remained at ceiling level for the other algorithms.
The sensitivity for RMA and IPD increased, but stayed below .51.
In a similar way, this supports the use of the PBF as a small sample size solution.
In this case when there are little available studies to synthesize evidence over.
It could be argued that when there are only 2 or 3 studies available to synthesize evidence over, that if a trade-off need be made, specificity is preferred over sensitivity.
It seems unwise to more easily acknowledge an effect if there is little data available.
However, the quality of the included studies should be considered before synthesizing evidence in any way.
Moreover, when the researcher performs some form of evidence synthesis, then the ability of the method to find an effect is a necessary and fair component, even if little studies are available.
Also, even with as much as ten available studies, PBF was the only algorithm that had a conventionally acceptable power.

@@

Overall, there was only a single marginal condition where the accuracy of PBF was lower than that of any other algorithm.
PBF scored 1-2 percentage points lower accuracy than IPD and RMA when there was perfect reliability.
Looking at the specificity and sensitivity however, this is explained by the relatively low specificity of PBF and the very high specificity of RMA and IPD.
Thus,
Within the scope of this simulation, PBF had relatively better inferential properties <!-- EB: higher accuracy --> than the other algorithms under consideration.
Although other algorithms had superior specificity, PBF had the highest levels of sensitivity.
Thus, PBF balanced the ability to accept an informative hypothesis in the presence of a true population effect with the ability to reject the hypothesis when there was no effect.

@@

As stated earlier, the PBF can be used as a small sample size solution,
although its increased type I error rate must be taken into account.
It implies that PBF works best as a complementary method for evidence synthesis, rather than a replacement.

@@

<!-- EB: I assume Rebecca means that we should note that 'the hypothesis being on average the best' is the hypothesis that reflects the truth. While 'the hypothesis that is best in all studies' is the hypothesis that is closest to the truth compared to the other hypotheses that were tested. It does not say anything necessarily about whether the tested hypotheses in general are good reflections of the truth. E.g, if we only test whether the earth is either flat-shaped or triangular and not acknowledge other options, then we would incorrectly conclude that the earth is flat.-->
The interpretation of the PBF should also be considered.
The PBF renders support for some hypothesis being the best in all studies, but only compared to other hypotheses it has been tested against.
This means that the tested hypothesis with most support is not necessarily the hypothesis that reflects the truth most closely.
This is because there are many other plausible hypotheses that have not been tested.
For example, testing whether the earth is flat or triangular excludes the possibility of a spherical earth.

@@

The argument from Rouder and Morey (2011) is that we cannot take the PBF as a valid measure for meta-analysis. They show that calculating the BF in multiple samples and taking their product results in a different PBF than when all data is pooled and the BF is calculated after pooling. They also show another way of calculating a meta-analytic BF and that this pooled BF comes closer to that meta-analytic BF. However, this calculation assume a constant population effect and the same type of statistical model across studies
. where all data is drawn from a common distribution The idiosyncratic differences between studies makes that pooling data is not a straightforward endeavour, because of the heterogeneity of the effect. Rouder and Morey do acknowledge their choice for a constant effect to be debatable.

@@

Rebecca noticed that Rouder and Morey specifically talk about PBF not being a good measurement in the context of testing H0, which is something that we are not doing. However, the only relevant difference I see in their vs our method is that their H0 contains a single equality constraint, while we used inequality constraints. Rouder and Morey do not explicitly point this difference out and I don't see how the PBF would be a bad or good idea depending on type of constraints.

They also do it for a single parameter t, but it would not work if there are multiple different units.-->

The article of @rouder2011bayes states that a meta-analytic Bayes factor is not as straightforward as taking the product of $k$ individual Bayes factors.
Rather, they provide a way of calculating more valid a meta-analytic Bayes factor.
However, only under the assumption that the population effect is constant and the test statistic is from the same type of statistical model.
Using this form of the PBF, they show that pooling the data and calculating a single Bayes factor on the pooled data comes closer to the meta-analytic Bayes factor than taking the product of each individual study.
However, as stated earlier, one strength of the PBF lies in its ability to synthesize evidence for a hypothesis, rather than an estimate.
This means that there is no assumption of a constant population effect and/or equal types of test statistics across studies.
Furthermore, as the article of @rouder2011bayes states itself, pooling data is valid if all data is independent and drawn from a common distribution.

@@

<!--EB: What I think Rebecca means here is the following: 
In classical NHST, there is an equality resitriction. we either do or do not reject H0, so either rho = x or rho != x. Rejecting H0, means that the data suggests that rho != .1, but rather something else than .1 (so .2, .9, -.9 etc.)
With RMA and IPD we effectively say, rho = .1 or rho != .1 which is true or not (even if the test statistic = -.8,than .1 is not in the 90% CI and we still H0, even though its in the opposite direction.)

@@

if rho = .1 and Rho = .1, than rejecting H0 is wrong. 
if rho = .2, and Rho = .1 than we must reject H0. -->

>>>>

This simulation study validated the product Bayes factor as an alternative or complementary method to synthesize evidence across studies.
The PBF was benchmarked against other common methods for evidence synthesis.
This was done by calculating a correlation coefficient and determining the accuracy, sensitivity and specificity as metrics of performance for the methods in different plausible varying conditions.
During the simulation, the hypotheses that were tested remained constant $H_i: \rho > .1$, $H_0: \rho = .1$.
To be able to calculate the performance metrics, half of the simulated conditions contained an effect, $\rho = .2$, while the other half did not, $\rho = .1$.
<!-- EB: Why did we not use H0:r=0 and Hi:r>0 as constant hypotheses? It seems more intuitive -->
Sensitivity and specificity can be interpreted as Type I error rate and statistical power respectively and are used interchangeably.
 <!-- Because Type 1 error rate and power seem to be key terms in the discussion, we may be able to include it in the introduction: 'How does the PBF compare to other state-of-the-art algorithms with regards to Type 1 error rate, power and overall accuracy?' -->
The effect of sample size per study, number of studies to synthesize evidence over and reliability of measurement on the performance of the algorithms was investigated.

@@

With an increasing number of groups, slightly decreasing specificity was traded off for increasing sensitivity.<!--EB: I would argue that the decrease is negligible, 0.767 at k = 2, 0.763 at k = 10 -->
With an increasing number of groups, specificity remained around .76-.77 for PBF, while sensitivity increased to .87.
Again, the specificity remained at ceiling level for the other algorithms.
The sensitivity for RMA and IPD increased, but stayed below .51.
In a similar way, this supports the use of the PBF as a small sample size solution.
In this case when there are little available studies to synthesize evidence over.
It could be argued that when there are only 2 or 3 studies available to synthesize evidence over, that if a trade-off need be made, specificity is preferred over sensitivity.
It seems unwise to more easily acknowledge an effect if there is little data available.
However, the quality of the included studies should be considered before synthesizing evidence in any way.
Moreover, when the researcher performs some form of evidence synthesis, then the ability of the method to find an effect is a necessary and fair component, even if little studies are available.
<!--EB: for clarification: I mean to say that research should not be done if there is effectively no power to detect an effect anyway, because the outcome is in such a case practically pre-determined. -->
Also, even with as much as ten available studies, PBF was the only algorithm that had a conventionally acceptable power.

@@

Overall, there was only a single marginal condition where the accuracy of PBF was lower than that of any other algorithm.
PBF scored 1-2 percentage points lower accuracy than IPD and RMA when there was perfect reliability.
Looking at the specificity and sensitivity however, this is mostly explained by the specificity at ceiling level of RMA and IPD.
Thus,
Within the scope of this simulation, PBF had relatively better inferential properties <!-- EB: higher accuracy --> than the other algorithms under consideration.
Although other algorithms had superior specificity, PBF had the highest levels of sensitivity.
Thus, PBF balanced the ability to accept an informative hypothesis in the presence of a true population effect with the ability to reject the hypothesis when there was no effect.

@@

Thus, the PBF can be used as a small sample size solution,
although its increased type I error rate must be taken into account.
It implies that PBF works best as a complementary method for evidence synthesis, rather than a replacement.

@@

<!-- EB: I assume Rebecca means that we should note that 'the hypothesis being on average the best' is the hypothesis that reflects the truth. While 'the hypothesis that is best in all studies' is the hypothesis that is closest to the truth compared to the other hypotheses that were tested. It does not say anything necessarily about whether the tested hypotheses in general are good reflections of the truth. E.g, if we only test whether the earth is either flat-shaped or triangular and not acknowledge other options, then we would incorrectly conclude that the earth is flat.-->
The interpretation of the PBF should also be considered.
The PBF renders support for some hypothesis being the best in all studies, but only compared to other hypotheses it has been tested against.
This means that the tested hypothesis with most support is not necessarily the hypothesis that reflects the truth most closely.
This is because there are many other plausible hypotheses that have not been tested.
For example, testing whether the earth is flat or triangular excludes the possibility of a spherical earth.
Still, because a Bayes factor is a continuous, real numbered quantity, it is more informative than the conclusion from a null-hypothesis significance test that only has the ability to dichotomously evaluate the null-hypothesis.
The main benefit is the diversity of hypotheses that can be tested using the PBF.

@@

The argument from Rouder and Morey (2011) is that we cannot take the PBF as a valid measure for meta-analysis. They show that calculating the BF in multiple samples and taking their product results in a different PBF than when all data is pooled and the BF is calculated after pooling. They also show another way of calculating a meta-analytic BF and that this pooled BF comes closer to that meta-analytic BF. However, this calculation assumes a constant population effect and the use of the same type of statistical model across studies. Also, pooling the data assumes that all data is drawn from a common distribution. The idiosyncratic differences between studies makes that pooling data is not a straightforward endeavour, because of the heterogeneity of the effect. Rouder and Morey do acknowledge their choice for a constant effect to be debatable.

@@

Rebecca noticed that Rouder and Morey specifically talk about PBF not being a good measurement in the context of testing H0, which is something that we are not doing. However, the only relevant difference I see in their vs our method is that their H0 contains a single equality constraint, while we used inequality constraints. Rouder and Morey do not explicitly point this difference out and I don't see how the PBF would be a bad or good idea depending on type of constraints. -->

One possible issue with the PBF as defined in this research comes from an article of @rouder2011bayes. 
The article states that a meta-analytic Bayes factor is not as straightforward as taking the product of $k$ individual Bayes factors.
<!--EB: Their reason seems to be that Bayes factors take sample size into account, effectively reducing the ability of a Bayes factor to recognize an effect in small sample. I dont see how this invalidates the PBF. If there is an effect, then in 5 small studies, it may be true that a BF is calculated that barely support the hypothesis of an effect. However, taking the product of 5 BFs that are all 1.25 will result in a PBF of about 3.-->
Rather, they provide a way of calculating more valid a meta-analytic Bayes factor.
However, only under the assumption that the population effect is constant and the test statistic is from the same type of statistical model.
Using this form of the PBF, they show that pooling the data and calculating a single Bayes factor on the pooled data comes closer to the meta-analytic Bayes factor than taking the product of each individual study.
However, as stated earlier, one strength of the PBF lies in its ability to synthesize evidence for a hypothesis, rather than an estimate.
This means that there is no assumption of a constant population effect and/or equal types of test statistics across studies.
Furthermore, as the article of @rouder2011bayes states itself, pooling data is preferred if all data is independent and drawn from a common distribution. 
The study idiosyncrasies encountered in evidence synthesis should wary the researcher in blindly making the assumption of a common distribution.

@@

<!--EB: What I think Rebecca means here is the following: 
In classical NHST, there is an equality restriction. we either do (r = .2) or do not (r = .1) reject H0. 
Rejecting H0 means that the data suggests that rho != .1, but rather something else than .1 (so .2, .9, -.9 etc.)
With RMA and IPD we effectively say, rho = .1 or rho != .1 which is true or not (even if the test statistic = -.8,than .1 is not in the 90% CI and we still reject  H0, even though the found effect is in the opposite direction of what we hypothesized.)

@@

This would imply that there should be more cases where the frequentist methods acknowledge an effect. However, we see a very low sensitivity due to the small effect, so its not really a concern in this paper-->
A final remark is that of performance comparison between PBF and the other algorithms which has to do with the differences in hypothesis specification. 
For the frequentist and PBF algorithms we specified $H_0: \rho = .1$ and $H_c: \rho \leq .1$ respectively.
This means, hypothetically, that if a strong effect in the opposite direction was found, let's say $\rho = -.6$, the frequentist algorithms would reject $H_0$ while the PBF would accept $H_c$.
This implies that there should have been a higher power to detect an effect in the frequentist framework, as negative effects would reject $H_0$.
This is contrary to what was observed in this research, however, as the frequentist approaches had a low sensitivity.
Again, this has to do with the small simulated difference between the true effect and the boundary value.
